\MakeUppercase{Graph Product Structure for Non-Minor-Closed Classes}

### 2022-05-21 k-planar.tex
fixed typo
added a little more to the map graph proof, and removed `spanning'

\begin{lem}
\label{MewMapGraphLemma}
For every integer $d\geq 3$, every $d$-map graph is a spanning \note{DW}{delete ``spanning''?} 
subgraph of a $d$-framed multigraph.
\end{lem}

@@

We may assume that $G_0$ is an edge-maximal multigraph with no faces of length 2 (and with each face labelled a nation or a lake), such that $G$ is the corresponding map graph. Suppose that some face $f$ of $G_0$ has a disconnected boundary (implying $G_0$ is disconnected). Let $v$ and $w$ be vertices in distinct components of the boundary of $f$. Add the edge $vw$ to $G_0$ across $f$. The corresponding map graph is unchanged, which contradicts the edge-maximality of $G_0$. Thus no face of $G_0$ is disconnected. Suppose that some face $f$ of $G_0$ has a repeated vertex in the boundary walk of $f$. Let $u,v,w$ be consecutive vertices on the boundary of $f$. So $u,v,w$ are distinct and $vw$ is not an edge in the boundary of $f$. Add the edge $uw$ across $f$ and consider the resulting face $uvw$ to be a lake. Since $v$ appears elsewhere in the boundary of $f$, the corresponding map graph is unchanged, which contradicts the edge-maximality of $G_0$. Thus no facial walk of $G_0$ has a repeated vertex. Since each facial walk is connected, every face of $G_0$ is bounded by a cycle, and $G_0$ is biconnected.

@@

By definition, $V(G)=V(\widehat{G}_1)$ \note{DW}{$V(\widehat{G}_1)=V(G_1)=V(G^*_0)$ which includes vertices corresponding to lakes of $G_0$. So I think this claim is wrong (or am I missing something?). I suggest we write $V(G) \subseteq V(\widehat{G}_1)$ and drop `spanning' from the lemma}. To prove the claim, it suffices to show that $E(G)\subseteq E(\widehat{G}_1)$.  Indeed, if $vw\in E(G)$ then the nation faces corresponding to $v$ and $w$ have a common vertex $x$ on their boundary. The vertex $x$ corresponds to a face $F_x$ in $G_0^*$ and the facial cycle of $F_x$ contains $v$ and $w$.  Therefore, the nation cycle $C_x$ of $F_x$ contains $v$ and $w$. If $C_x$ has length $2$ then $vw\in E(G_1)\subseteq E(\widehat{G}_1)$.  If $C_x$ has length at least $3$ then $C_x$ bounds a face in $G_1$, so $vw\in E(\widehat{G}_1)$.
\end{proof}

>>>>

\begin{lem}
\label{MewMapGraphLemma}
For every integer $d\geq 3$, every $d$-map graph is a subgraph of a $d$-framed multigraph.
\end{lem}

@@

If $G_0$ has a face $f$ of length 2, then add a new vertex inside $f$ adjacent to both vertices on the boundary of $f$, which creates two new triangular faces $f_1$ and $f_2$. If $f$ is a lake, then make $f_1$ and $f_2$ lakes. If $F$ is a nation, then make $f_1$ a nation and make $f_2$ a lake. The resulting map graph is still $M$. So we may assume that $G_0$ is an edge-maximal multigraph with no faces of length 2 (and with each face labelled a nation or a lake), such that $G$ is the corresponding map graph. Suppose that some face $f$ of $G_0$ has a disconnected boundary (implying $G_0$ is disconnected). Let $v$ and $w$ be vertices in distinct components of the boundary of $f$. Add the edge $vw$ to $G_0$ across $f$. The corresponding map graph is unchanged, which contradicts the edge-maximality of $G_0$. Thus no face of $G_0$ is disconnected. Suppose that some face $f$ of $G_0$ has a repeated vertex in the boundary walk of $f$. Let $u,v,w$ be consecutive vertices on the boundary of $f$. So $u,v,w$ are distinct and $uw$ is not an edge in the boundary of $f$. Add the edge $uw$ across $f$ and consider the resulting face $uvw$ to be a lake. Since $v$ appears elsewhere in the boundary of $f$, the corresponding map graph is unchanged, which contradicts the edge-maximality of $G_0$. Thus no facial walk of $G_0$ has a repeated vertex. Since each facial walk is connected, every face of $G_0$ is bounded by a cycle, and $G_0$ is biconnected.

@@

By definition, $V(G) \subseteq V(\widehat{G}_1)$. To prove the claim, it suffices to show that $E(G)\subseteq E(\widehat{G}_1)$.  Indeed, if $vw\in E(G)$ then the nation faces corresponding to $v$ and $w$ have a common vertex $x$ on their boundary. The vertex $x$ corresponds to a face $F_x$ in $G_0^*$ and the facial cycle of $F_x$ contains $v$ and $w$.  Therefore, the nation cycle $C_x$ of $F_x$ contains $v$ and $w$. If $C_x$ has length $2$ then $vw\in E(G_1)\subseteq E(\widehat{G}_1)$.  If $C_x$ has length at least $3$ then $C_x$ bounds a face in $G_1$, so $vw\in E(\widehat{G}_1)$.
\end{proof}
